---
title: "Making an antigenic map from titer data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making an antigenic map from titer data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we're going to go through the process of making an antigenic map using Racmacs and viewing and saving the resulting map and images and interactive plots of the map.

The steps will be:

- Reading in the titer data
- Creating the acmap object
- Running a set of optimizations to try and find the best map to represent the data
- Viewing the map
- Customising map
- Saving the map

## Reading in the titer data
First of all we need to read in the titer data from a file, exactly how you do this depends on the format of your data but ultimately you need to end up with a matrix or dataframe of titers where _rows represent antigens_ and _columns represent sera_.

If your titer data is in a simple format where you just have a single header for serum names and a single first column of antigen names you can try using the function `read.titerTable()`, as in the example below.

For this example we will be reading the titer data from the first antigenic map of H3N2 antigenic evolution, published in 2004 and included with the Racmacs package.

```{r reading_titers}
# Load the Racmacs package
library(Racmacs)

# Read in the titer table
path_to_titer_file <- system.file("extdata/h3map2004_hitable.csv", package = "Racmacs")
titer_table        <- read.titerTable(path_to_titer_file)
```

The resulting titer table is a matrix of titers with column and row names:
```{r}
print(titer_table[1:5,1:7])
```

## Creating the acmap object
Now we have the titers in the correct format we can use it create an antigenic map object or "acmap" object. For this we use the constructor function `acmap()`, akin to how `data.frame()` is used to make a data frame object or `matrix()` for a matrix.

```{r making_an_acmap}
# Create the acmap object, specifying the titer table
map <- acmap(
  titer_table = titer_table
)
```

The resulting object, that we've given the name `map` in this case, only has properties associated with the table. If the table had column names and row names, serum and antigen names are inferred from these. Alternatively, you can set these explicitely when creating the map with the `ag_names` and `sr_names` arguments.

## Running optimizations
Although we now have the acmap object there is no "map" yet as such, since we need to invoke the optimiser function to try and determine the antigenic map that best represents the patterns of reactivity seen in the titer table. To do this we call the function `optimizeMap()` on the acmap object we've made, specifying the number of optimization runs and other parameters like the minimum column basis and number of dimensions.

In this case we're going to try and find the best 2-dimensional solution for the map data, using 500 optimization runs.

```{r optimizing_an_acmap}
# Perform some optimization runs on the map object to try and determine a best map
map <- optimizeMap(
  map                     = map,
  number_of_dimensions    = 2,
  number_of_optimizations = 500,
  minimum_column_basis    = "none"
)
```

By default the resulting map optimizations are sorted by their stress, so that the optimization that resulted in the lowest map stress (i.e. usually thought of as the best solution), will be first and will be the default optimization run selected and therefore will be the one used when querying anything like antigen and serum coordinates, map stress etc. unless otherwise specified.

## Viewing the map
The lowest stress optimization run will also be the one used by default when viewing the map. To view the map, you can either do a simple plot of the map using the `plot()` function

```{r plotting_an_acmap, fig.width=10, fig.height=6, out.width="100%", fig.retina=TRUE}
plot(map)
```

or you can view the map interactively using the `view()` function

```{r viewing_an_acmap, out.width = '100%'}
view(map)
```

### Using `make.acmap()`
Note that there is also a small utility function provided to allow you to create the acmap object and run optimizations on it all in one step. In the example below we use it to create a 3d version of the antigenic map and then use `view()` map to view the result (`plot()` will not work with 3 dimensional maps).

```{r making_a_3d_map, out.width = '100%'}
# Make the acmap object and run optimizations
map3d <- make.acmap(
  titer_table             = titer_table,
  number_of_dimensions    = 3,
  number_of_optimizations = 500,
  minimum_column_basis    = "none"
)

# View the result
view(map3d)
```

## Customising the map

### Adding antigen and sera names

```{r adding_names}
# Add sera and antigen names
srNames(map) <- c(colnames(titer_table))
agNames(map) <- c(rownames(titer_table))
```
When viewing the map using `view()`, the names will appear upon hovering over antigen or sera points, or when hitting `n` on the keyboard.

You can add antigen names to a static plot like so:

```{r add_names_to_static_plot, fig.width=10, fig.height=6, out.width="100%", fig.retina=TRUE}
plot(
  map,
  # grid.col = "#666666" to change the colour of the grid
)

text(
  agCoords(map),
  labels = agNames(map)
)
```
To add antigen names to an interactive plot use:

```{r add_names_to_interactive_plot, out.width = '100%'}
map <- r3js::text3js(
  map,
  x = agCoords(map)[,1],
  y = agCoords(map)[,2],
  z = rep(0, numAntigens(map)),
  text = agNames(map),
  type = "html",
  # offset = c(10,0),    # You can play around with offset and alignment to get text to appear e.g. at the right hand side
  alignment = "right"
)

view(
  map
  # options = list(
  #   grid.col = "#666666"  to change the colour of the grid
  # )
)
```

### Changing look of antigen and sera

You can customise the colour, outline width and colour, and size of antigens using the `agFill()`, `agOutlineWidth()`, `agOutline()`, and `agSize()` functions. Sera can be changed analogously using `srFill()`, `srOutlineWidth()`, `srOutline()`, and `srSize()`.

```{r setting_cols_etc, out.width = '100%'}
# Change fill color or outline color of specific antigen by name or index
agFill(map)[agNames(map) == "BI/2668/70"] <- "cyan"
agOutline(map)[10] <- "orange"
agOutlineWidth(map)[10] <- 3

# Or change using logical vector to select
blue_ags <- agNames(map) %in% c("HK/3/95", "HK/32/95")
agFill(map)[blue_ags] <- "blue"

# Plot or view interactively or export interactive viewer in webpage
view(map)
```

You can also add the point styles from a previously made map to your current map:

```{r add_plotspec}
# Read previously made map:
oldMap <- read.acmap("extdata/h3map2004_bootstrap.ace")

map <- applyPlotspec(map, oldMap)

view(map)
```

## Saving a map

Once the map has been optimised and customised, you can either save the entire `acmap` data object, or just the antigen and sera coordinates:

```{R eval = FALSE}
# save the whole acmap data object:
save.acmap(map, filename)

# save the coordinates of antigens and sera:
save.coords(
  map,
  filename,
  optimization_number = 1,
  antigens = TRUE,
  sera = TRUE
)
```
